Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning

### 2023-02-07 paper/fig/arh_binomial.tex
impact of budgets on coverage set

\begin{tikzpicture}
\pgfplotsset{
    scale only axis,
    % height=6cm,
    xlabel={total daily new hospitalizations},
    ylabel={total lost contacts},
    width=0.45\textwidth
}
\begin{groupplot}[
    group style={
        group size=2 by 1,
        y descriptions at=edge left,
        vertical sep=0pt,
        horizontal sep=0pt,
    },
    legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
]
\nextgroupplot[mark size=1pt,draw opacity=1]
        \addplot[only marks,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_2.csv};
        \addplot[only marks,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_3.csv};
        \addplot[only marks,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_4.csv};
        \addplot[only marks,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_5.csv};
        \addplot[only marks,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/ode/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
\nextgroupplot[mark size=1pt]
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_3.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_4.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_5.csv};
        \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_None.csv};
        \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};

\end{groupplot}
\end{tikzpicture}

>>>>

\begin{tikzpicture}
    \pgfplotsset{
        scale only axis,
        % height=6cm,
        xlabel={total number of daily new hospitalizations},
        ylabel={total lost contacts}
    }
    \begin{axis}[
    legend entries={{$\text{PCN}_{b=2}$}, {$\text{PCN}_{b=3}$}, {$\text{PCN}_{b=4}$}, {$\text{PCN}_{b=5}$}, {$\text{PCN}$}, {Fixed}},
    % width=9.5cm,
    mark size=1pt,
    ]
    \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc1] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_2.csv};
    \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc2] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_3.csv};
    \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc3] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_4.csv};
    \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc4] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_5.csv};
    \addplot[only marks,draw opacity=1,fill opacity=0.5,mark=o,pc5] table [x=o_1, y=o_5, col sep=comma] {data/cs/arh/binomial/budget_None.csv};
    \addplot[only marks,mark=+] table [x=o_0, y=o_1, col sep=comma] {data/cs/cs_fixed.csv};
    \end{axis}
\end{tikzpicture}

### 2023-02-07 paper/main.tex
impact of budgets on coverage set

\newcommand{\foi}{\lambda}
\newcommand\added[1]{\textcolor{blue}{#1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysing COVID-19 deconfinement policies}
\label{sec:experiments}
Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment. We aim to obtain policies that balance between the social burden experienced by the population and the epidemiological objective of minimising the attack rate. To this end we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.

We now evaluate our extension of PCN for continuous action-spaces on both the ODE and Binomial models. Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population. Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown of the country. This is implemented by fixing the actions $p_w, p_s, p_l$ to $0.2, 0, 0.1$ respectively. This lockdown ended on the 4th of May 2020, at which point the government decided on a multi-phase exit strategy to incrementally reduce teleworking, reopen schools and allow leisure activities, such as the re-opening of bars and the cultural sector. It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily new hospitalizations and the total number of contacts lost as a proxy for social burden. The simulation lasts throughout the school holidays (from 01/07/2020 to 31/08/2020). Schools are closed during the school holidays, which is simulated by setting $p_s = 0$, regardless of the corresponding value outputted by the policy, i.e., during periods of school closure $p_s$ is ignored.

To evaluate the quality of the policies discovered by PCN, we compare it to a baseline. The baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging from 0 to 1. In other words, the fixed policies directly operate in a fine-grained manner on the whole contact reduction function $\hat{C}$. This allows us to obtain a strong baseline for potential exit strategies over the objective space. We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is what we aim to provide using PCN.

@@

\begin{figure}
    \centering
    \input{fig/arh.tex}
    \caption{The Pareto front of policies discovered by PCN using the ODE model (Binomial omitted for clarity, see SI), showing the different compromises between the number of hospitalizations and the number of lost contacts. Although $\text{PCN}_\text{ARI}$ was trained on the number of infections, it still shows a competitive coverage set with respect to hospitalizations.}
    \label{fig:pareto-front}
\end{figure}

@@

We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}). The coverage set is shown in Figure~\ref{fig:pareto-front}.

We notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, showing that there are much better alternatives to the fixed policies. This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations. In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g. more than $20000$) for the same social burden. Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.

@@

\begin{figure}
    \centering
    \input{fig/all_budgets}
    \caption{Selection of policies learned by PCN, from most restrictive in terms of social burden (top) to least restrictive (bottom).
    }
    \label{fig:policy-execution}
\end{figure}

>>>>

\newcommand{\foi}{\lambda}
\newcommand\added[1]{\textcolor{blue}{#1}}
\newcommand\todo[1]{\textcolor{red}{TODO: #1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysing COVID-19 deconfinement policies}
\label{sec:experiments}
Our goal is to use PCN to learn deconfinment strategies in the \momdpname\ environment. We aim to learn policies that balance between the epidemiological objective of minimising the attack rate and the social burden experienced by the population. To this end we consider two cases for the vectorial reward functions $[\mdprewardfn_\text{ARH}, \mdprewardfn_\text{SB}]$ and $[\mdprewardfn_\text{ARI}, \mdprewardfn_\text{SB}]$, to learn and analyse policies under different targets with respect to the considered attack rate.

To conduct this analysis, we evaluate our extension of PCN for continuous action-spaces on both the Binomial model. Conform to \citet{abrams2021modelling}, the simulation starts on the 1st of March 2020, by seeding a number of infections in the population. Two weeks later, on the 14th of March, the Belgian government initiated a full lockdown of the country. This is implemented by fixing the actions $p_w, p_s, p_l$ to $0.2, 0, 0.1$ respectively. This lockdown ended on the 4th of May 2020, at which point the government decided on a multi-phase exit strategy to incrementally reduce teleworking, reopen schools and allow leisure activities, such as the re-opening of bars and the cultural sector. It is from this day onward that PCN aims to learn policies for diverse exit strategies, compromising between the total number of daily new hospitalizations and the total number of contacts lost as a proxy for social burden. The simulation lasts throughout the school holidays (from 01/07/2020 to 31/08/2020). Schools are closed during the school holidays, which is simulated by setting $p_s = 0$, regardless of the corresponding value outputted by the policy, i.e., during periods of school closure $p_s$ is ignored.

% https://nl.wikipedia.org/wiki/Maatregelen_tijdens_de_coronacrisis_in_Belgi%C3%AB
We draw the analogy with the multi-phase exit strategy established by the Belgian government and the restriction on the number of action-changes imposed by the \momdpname's budget $\budget$. Indeed, on the 11th of May 2020, exactly one week after the end of the lockdown, stores and certain types of jobs were allowed to reopen, under strict conditions. This corresponds to altering $p_w$ in our MOMDP. One week later, on 18th of May, primary and secondary schools reopened for limited sized class-groups, and the cultural sector reopened partially. This is equivalent to increasing $p_s$ and $p_l$. Further changes of restrictions occured on the 8th of June, 1st, 9th and 25th of July. Thus, we argue that, with a limited budget, we can achieve realistic and implementable policies. In our experiments, we consider budgets of 2 to 5, as these closely relate to the number of changes that occured until the end of the summer holidays of 2020. As an upper bound, we also consider a no-limit budget setting.

To evaluate the quality of the policies discovered by PCN, we compare it to a baseline. The baseline consists of a set of 100 fixed policies, that iterate over all the possible social restriction levels, with values ranging from 0 to 1. Concretely, each policy uses a fixed proportional reduction $p_w=p_l=p_s=u, u\sim\mathcal{U}(0, 1)$ throughout the episode. In other words, the fixed policies directly operate in a fine-grained manner on the whole contact reduction function $\hat{C}$. This allows us to obtain a strong baseline for potential exit strategies over the objective space. We note that while such fixed policies are a feasible approach, they do not scale well in terms of action and objective spaces and they will not be able to provide an adaptive restriction level, which is what we aim to provide using PCN.

@@

\begin{figure}
    \centering
    \input{fig/arh_binomial.tex}
    \caption{The Pareto front of policies discovered by PCN using the Binomial model, showing the different compromises between the number of hospitalizations and the number of lost contacts. As the budget increases, so does the coverage set learnt by PCN. Changes are most noticeable in the less restrictive trade-offs in terms of social burden.}
    \label{fig:pf-binomal-arh-budgets}
\end{figure}

@@

We learn a coverage set that ranges from imposing minimal restrictions to enforcing many restrictions (see Figure~\ref{fig:policy-execution}). The coverage set is shown in Figure~\ref{fig:pf-binomal-arh-budgets}.

Regardless of the imposed budget, we notice that the coverage sets discovered by PCN almost completely dominate the coverage set of the baseline, demonstrating that there are better alternatives to the fixed policies. This is most visible in the compromising policies, where one has to carefully choose when to remove social restrictions while at the same time minimizing the impact on daily new hospitalizations. In these scenarios, PCN discovers policies that can drastically reduce the total number of new hospitalizations (e.g. more than $20000$) for the same social burden. Analysing the executions of such policies (see Figure~\ref{fig:policy-execution}, middle plot) shows a flattened hospitalization curve, with a gradual increase of social freedom during the school holidays such that the peak stays stable and gradually decreases over time.

@@

\subsection{Impact of budgets on the coverage set}

Looking at the coverage sets displayed in Figure~\ref{fig:pf-binomal-arh-budgets}, we notice that the budget impacts the coverage set that has been learnt. In general, an increase of budget results in an increasingly better coverage set. This is to be expected, as a higher budget give the agent more freedom to change its actions as the epidemic progresses. However, we also notice that this increase of coverage sets is not necessarily gradual. Between the run with a budget of 2 and budget of 3, the difference in coverage set seems marginal. \todo{insight}.

Moreover, we observe that the improvement concentrates around the less restrictive policies in terms of social burden. We believe that this regio contains the most complex policies, as these try to maintain as much social freedom as possible, while containing the number of hospitalisations. We believe that, in these cases, the timing of the actions greatly correlate with the timing and duration of the peak of the epidemic, and a higher budget allows for more fine-grained control about managing this timing. To confirm this, we select, for each budget setting, the solution where the difference in performance is most noticeable, corresponding to the solutions with a total number of hospitalisations around $80000$. We plot the execution of the corresponding policies in Figure~\ref{fig:pe-binomial-arh-budgets} and analyse the difference they make in terms of social burden. First, we notice that the lower-budget policies are unable to lessen the social restrictions past the peak of the epidemic. In contrast, the setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave. We note how the policy with a budget of 5 ressembles the execution of the one without restrictions but that, due to its budget, is unable to progressively lessen the restrictions and instead resorts to a halfway compromise. \todo{corresponding SB numbers}

Compared to this specific regio of the coverage set, the difference in performance between different budget settings seems marginal around the extremas. \todo{explain}

Finally, we observe that, while the extremas deliver similar trade-offs regardless of the chosen budget, the results differ for the setting without budget restrictions. Indeed, in this setting, PCN does not seem to learn the most and least restrictive policies, even though there are no constraints on the action-set. As explained in Section~\ref{sec:pcn-policy-exploration}, PCN searches for increasingly better solutions using a stochastic policy. Thus, at every timestep, the action can change compared to the previous one. Continuously outputting the same action (e.g., no social restrictions) becomes a complicated task. In comparison, for the settings with a limited budget, the action stays the same as the previous one once the budget has been spent. It is thus easier to discover the most extreme policies. Thus, in the specific case where we impose no budget, we believe the freedom of action actually hinders PCN's search.

\begin{figure}
    \centering
    \input{fig/all_budgets}
    \caption{Execution of the policies attaining a number of hospitalisations around $80000$, for different budgets. From top to bottom are displayed the policy executions of runs with budget 2,3,4,5 and no-limit, respectively. We notice that the lower-budget policies are unable to lessen the social restrictions past the peak. The setting with no budget restrictions finely controls the restrictions as the epidemic progresses, completely removing restrictions by the end of the wave. Finally, there is no consensus on which social environment to restrict most, some policies providing similar restrictions for $p_w$ and $p_l$, while others putting harsher restrictions on one social environment than the other.
    }
    \label{fig:pe-binomial-arh-budgets}
\end{figure}
